Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-17 article/draft.tex
few typos in intro + intro of section 1, EV

\section*{Abstract}

@@

@@

During the COVID-19 global health pandemic platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see Appendix \ref{links} for references). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report}''. In particular, policies regarding misinformation are generally not part of the set of platform rules or community guidelines (as of July 2021).  %Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

Misinformation related interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify with data mining mainstream platforms' current actions regarding content with low credibility or false information. We do so by providing a series of examples for different interventions and platforms. We chose to focus on three platforms: Facebook, Twitter and YouTube. Both Facebook and YouTube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. To collect data from these three platforms, we either used the APIs (Application Programming Interfaces), or web scraping, i.e. retro-engineering the HMTL code of a web page to extract meaningful data, see Table \ref{tab1} for a summary. Minet \cite{minet}, a webmining tool developed by the Médialab SciencesPo, was often used, and we scripted our own data mining code when it was necessary.

@@

More specifically, we survey in this article a number of common policies against misinformation used by Facebook, Twitter and YouTube. We classified those common policies into three broad categories: $(i)$ temporary or permanent suspension of users, $(ii)$ introducing flags and notices, and $(iii)$ reducing the visibility of some content. We compile in Appendix \ref{links} in table \ref{summary}, a list of links that redirect to the policies, regulations and transparency centers of Facebook, Twitter and YouTube that we discuss throughout the article. Furthermore, the examples provided to illustrate how to monitor a given policy were picked out of a list of domain names with several failed fact-checks. To be more precise, some domain names with failed fact-checks were picked because a given platform communicated about an intervention or because an intervention (e.g. suspension) was announced on the social media accounts linked to a given domain name. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.

@@

Mainstream social media platforms may suspend the account of a specific user when they deem that the platforms' rules have been violated. Account suspension can be temporary or permanent.  When the suspension is temporary the user is prohibited for a limited period of time from posting content on their account, but created content prior to suspension remains available to the user and their followers. However, when the suspension is permanent, in most cases, followers or subscribers have no longer access to the content prior to the suspension and the user can no longer use the account to create new content.  In what follows, we focus on the implementation of this policy by Facebook, Twitter and YouTube, and we provide simple examples to illustrate.

>>>>

\begin{abstract}

@@

\end{abstract}

@@

During the COVID-19 global health pandemic, platforms have upgraded their guidelines to include a set of rules to tackle the propagation of potentially harmful content (see Appendix \ref{links} for references). Those policies are enforced via existing actions used by the platforms to tackle other rules' violations, such as: labelling content to provide more context or indicate falsehood, publishing a list of terms or topics that will be flagged, suspending accounts, implementing strike systems, reducing the visibility of content, etc. As each platform is a private company, those {\it new} policies are not coordinated and are implemented in different ways across platforms. Such targeted policies show the willingness of mainstream platforms to enhance the quality of the online conversation, but also sheds light on the lack of specific policies to tackle misinformation in general. The 2019 report of the Facebook Data Transparency Advisory Group (DTAG) states that ``{\it DTAG was not tasked with evaluating any of the following: (...) Facebook’s policies with respect to “fake news” or misinformation, as neither of these categories were counted as violations within the first two versions of the Community Standards Enforcement Report}''. In particular, policies regarding misinformation are generally not part of the set of platform rules or community guidelines (as of July 2021).  %Facebook's strategy to tackle ``False News'' is three fold : Remove, Reduce, Inform. It is explained via a blog post on the facebook Newsroom. Similarly, Twitter communicates about actions related to misinformation via their \href{https://blog.twitter.com/en\_us/authors.TwitterSafety}{Twitter Safety} Blog/account. In a post on YouTube Official Blog, the platform explained its ``Four Rs of Responsibility" and how it raises authoritative content, reduces borderline content and harmful misinformation (see table \ref{summary} in Appendix \ref{links}).

@@

Misinformation related interventions by mainstream platforms are hard to monitor, study or verify by third parties (e.g. academic community, data journalists, NGOs), as in many cases misleading or false content does not qualify as a violation of a given platform's rules and it does not explicitly appear as a separate category in available transparency reports. This makes the study of online misinformation, the assessment of the impact of platforms' actions to tackle misinformation and their relevancy a burdensome task for the academic community. Hence, in the present article we explain how to verify 
mainstream platforms' current actions regarding content with low credibility or false information
with data mining. We do so by providing a series of examples for different interventions and platforms. We chose to focus on three platforms: Facebook, Twitter and YouTube. Both Facebook and YouTube are in the top three most popular social media platforms in terms of number of users.\footnote{See for example the ranking of the most popular social networks as of April 2021 on Statista: \href{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}{https://www.statista.com/statistics/272014/global-social-networks-ranked-by-number-of-users/}.} We further choose Twitter because it is a social networking platform with the most news-focused users, according to the Pew Research center (2019)~\cite{pew1}. To collect data from these three platforms, we either used the APIs (Application Programming Interfaces), or web scraping, i.e. retro-engineering the HMTL code of a web page to extract meaningful data, see Table \ref{tab1} for a summary. Minet \cite{minet}, a webmining tool developed by the SciencesPo médialab, was often used, and we scripted our own data mining code when it was necessary.

@@

More specifically, we survey in this article a number of common policies against misinformation used by Facebook, Twitter and YouTube. We classified those common policies into three broad categories: $(i)$ temporary or permanent suspension of users, $(ii)$ 
informing users with flags and notices, and $(iii)$ reducing the visibility of some content. We compile in Appendix \ref{links} in table \ref{summary}, a list of links that redirect to the policies, regulations and transparency centers of Facebook, Twitter and YouTube that we discuss throughout the article. Furthermore, the examples provided to illustrate how to monitor a given policy were picked out of a list of domain names with several failed fact-checks. To be more precise, some domain names with failed fact-checks were picked because a given platform communicated about an intervention or because an intervention (e.g. suspension) was announced on the social media accounts linked to a given domain name. Finally, we discuss how an increased effort of transparency regarding specific content can help the community of researchers study and assess the impact of platforms' policies regarding misinformation.

@@

Mainstream social media platforms may suspend the account of a specific user when they deem that the platforms' rules have been violated. Account suspension can be temporary or permanent.  When the suspension is temporary the user is prohibited for a limited period of time from posting content on their account, but 
content created prior to suspension remains available to the user and their followers. However, when the suspension is permanent, in most cases, followers or subscribers 
no longer have access to the content prior to the suspension and the user can no longer use the account to create new content. In what follows, we focus on the implementation of this policy by Facebook, Twitter and YouTube, and we provide simple examples to illustrate.
